Modelling temporal dependence and multilevel structure using hidden Markov models

### 2020-07-22 Report/Sections/simulation.tex
Updates from Overleaf

The HHMM-DFT performed almost identically to the CarHHMM-DFT in most respects, including its decoding accuracy and parameter bias. 
%- it had very low error for the fine-scale hidden states and $\approx 94\%$ accuracy on the dive types (see Table \ref{table:accuracy}). In addition, the standard errors of the parameter estimates were comparable to those of the CarHHMM-DFT. The estimates of standard error using the observed Fisher information were also comparable to the CarHHMM-DFT. 
The HHMM-DFT \textit{did} differ from the CarHHMM-DFT in its estimate of the standard deviation of acceleration, $\hat \sigma_1^*$. For both sub-dive states, $\hat \sigma_1^*$ greatly overestimated the true standard deviation. In particular, $\bar \hat \sigma_1^{*(1)} = 0.248$ and $\bar \hat \sigma_1^{*(2)} = 0.242$, but $\sigma_1^{*(1)} = 0.05$ and $\sigma_1^{*(2)} = 0.1$. in addition, the empirical standard error of both $\hat \mu_1^*$ and $\hat \sigma_1^*$ was much larger than the estimates of standard error using the observed fisher information (see Section S-1.2, Table 3). This implies that parameter uncertainty can be underestimated when auto-correlation is ignored.

The CarHHMM was the worst-performing model in terms of decoding accuracy, with an accuracy of $89 \pm 2\%$ when determining sub-dive behavior and an accuracy of $91 \pm 3\%$ when determining dive types. It was much more likely to misclassify sub-dive behavior 1 than sub-dive behavior 2 (see Section S-1.2, Table 1). It was also the worst model in terms of parameter estimates, in part due to its relatively poor decoding accuracy. $\hat \Theta^{*(2)}$ and $\hat \Gamma^{*(2)}$ in particular were poor estimates of the true parameters (see Section S-1.1.3).

Finally, the CarHMM-DFT was comparable to the CarHHMM-DFT in terms of decoding accuracy, parameter bias, and parameter standard error for the fine-scale observations. In addition, the time required to fit the CarHHMM was approximately 1/5 that of the other models. However, it lacked any hierarchical structure, so it was unable to decode the coarse-scale dive types. The CarHMM-DFT still did fit a gamma distribution to the dive duration of \textit{all} dives. The parameter estimates $\hat \mu$ and $\hat \sigma$ are highly correlated, and $\hat \sigma$ is a significant underestimate by $\approx$ 5 seconds.

>>>>

The HHMM-DFT performed almost identically to the CarHHMM-DFT in most respects, including its decoding accuracy and parameter bias. 
%- it had very low error for the fine-scale hidden states and $\approx 94\%$ accuracy on the dive types (see Table \ref{table:accuracy}). In addition, the standard errors of the parameter estimates were comparable to those of the CarHHMM-DFT. The estimates of standard error using the observed Fisher information were also comparable to the CarHHMM-DFT. 
The HHMM-DFT \textit{did} differ from the CarHHMM-DFT in its estimate of the standard deviation of acceleration, $\hat \sigma_1^*$. For both sub-dive states, $\hat \sigma_1^*$ greatly overestimated the true standard deviation. In particular, $\overline{\hat \sigma_1^{*(1)}} = 0.248$ and $\overline{\hat \sigma_1^{*(2)}} = 0.242$, but $\sigma_1^{*(1)} = 0.05$ and $\sigma_1^{*(2)} = 0.1$. In addition, the empirical standard error of both $\hat \mu_1^*$ and $\hat \sigma_1^*$ was much larger than the estimates of standard error using the observed fisher information (see Section S-1.2, Table 3). This implies that parameter uncertainty can be underestimated when auto-correlation is ignored.

The CarHHMM was the worst-performing model in terms of decoding accuracy, with an accuracy of $89 \pm 1\%$ when determining sub-dive behavior and an accuracy of $87 \pm 11\%$ when determining dive types. It was much more likely to misclassify sub-dive behavior 1 than sub-dive behavior 2 (see Section S-1.2, Table 1). It was also the worst model in terms of parameter estimates, in part due to its relatively poor decoding accuracy. $\hat \Theta^{*(2)}$ and $\hat \Gamma^{*(2)}$ in particular were poor estimates of the true parameters (see Section S-1.1.3).

Finally, the CarHMM-DFT was comparable to the CarHHMM-DFT in terms of decoding accuracy, parameter bias, and parameter standard error for the fine-scale observations. In addition, the time required to fit the CarHHMM was less than half of that of the other models (Tbl \ref{table:accuracy}). However, it lacked any hierarchical structure, so it was unable to decode the coarse-scale dive types. The CarHMM-DFT still did fit a gamma distribution to the dive duration of \textit{all} dives. The parameter estimates $\hat \mu$ and $\hat \sigma$ are highly correlated, and $\hat \sigma$ underestimates the true standard deviation of all dive types by $\approx$ 5 seconds.
